R for Clinical Study Reports and Submission

### 2021-09-15 tlf-overview.Rmd
init commit

Within eCTD, clinical study reports (CSR) are located at module 5. 
[ICH E3 guidance](https://www.ich.org/page/efficacy-guidelines) provide 
compilation of structure and content of clinical study reports.

@@

Building CSR is a team work with clinicians, medical writers, statisticians and statistical programmers.
Here, we focus on the work and deliverables completed by statisticians and statistical programmers. 
In an organization, A group of statisticians and statistical programmers commonly work together to 
define, develop, validate and deliver tables, listings and figures (TLFs) required for a CSR.
Microsoft Word is widely used to prepare CSR in pharmaceutical industry. 
Therefore, `rtf`, `doc`, `docx` are commonly used format to deliver TLFs.

>>>>

Within eCTD, clinical study reports (CSR) are located at module 5. 
[ICH E3 guidance](https://www.ich.org/page/efficacy-guidelines) provides 
compilation of structure and content of clinical study reports.

@@

Building CSR is a team work with clinicians, medical writers, statisticians and statistical programmers.
Here, we focus on the work and deliverables completed by statisticians and statistical programmers. 
In an organization, a group of statisticians and statistical programmers commonly work together to 
define, develop, validate and deliver tables, listings and figures (TLFs) required for a CSR.
Microsoft Word is widely used to prepare CSR in pharmaceutical industry. 
Therefore, `rtf`, `doc`, `docx` are commonly used format to deliver TLFs.

### 2021-09-15 tlf-population.Rmd
init commit

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 11.1, data sets analysed of a CSR.

@@

As discussed in 
[R for data science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function 
whenever you’ve copied and pasted a block of code more than twice"

@@

- format percentage using `formatC` 
- calculate number and percentage by groups

@@

To format number and percentage, we created a function called `fmt_num`.
It is a very simple function that is a wrapper of `formatC`.

@@

To calculate number and percentage of patients by groups, 
we provide a simple (but not robust) wrapper function, `count_by` using `dplyr` and `tidyr`.

@@

Now we combined individual rows into the whole table for reporting purpose. 
`tbl_pop` is used as input for `r2rtf` to create final report.

@@

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in Section 11.1, Data Sets Analysed.

@@

As discussed in 
[R for data science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function 
whenever you’ve copied and pasted a block of code more than twice".

@@

- format percentages using `formatC` 
- calculate numbers and percentages by groups

@@

To format numbers and percentages, we created a function called `fmt_num`.
It is a very simple function wrapping `formatC`.

@@

To calculate numbers and percentages of patients by groups, 
we provide a simple (but not robust) wrapper function, `count_by` using `dplyr` and `tidyr`.

@@

Now we combine individual rows into the whole table for reporting purpose. 
`tbl_pop` is used as input for `r2rtf` to create final report.

@@

In conclusion, the procedure to generate a population summary table as shown in the above example is listed as follows:

- Step 1: Read data into R, i.e.,  `adsl`.
- Step 2: Rowly bind the counts/percentages of the ITT population, the efficacy population and the safety population. These these types of population are generated by a help function `count_by`. 
- Step 3: Format the output in Step 2 by `r2rtf`.
